---
title: "STA141_Project"
author: "Trevor Carpenter, Ryan Buchner, Bobby Chow"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
bank_data = read_delim("bank-additional/bank-additional.csv", ";", col_names=TRUE, col_types = "dfffffffffddddfdddddf")
```

First we process the data. Education has a logical order to it so we can convert it to a numeric. This works for trees, but not necessarily for neural networks or logistic since the distances between the levels which we made all equal to 1 are arbitrary.
```{r}
## process data for education
bank_data$education = as.numeric(ifelse(bank_data$education == "illiterate", 0, ifelse(bank_data$education == "basic.4y", 1, ifelse(bank_data$education == "basic.6y", 2, ifelse(bank_data$education == "basic.9y", 3, ifelse(bank_data$education == "high.school", 4, ifelse(bank_data$education == "university.degree", 5, ifelse(bank_data$education == "professional.course", 0, NA))))))))

## omit na values
bank_data = na.exclude(bank_data)
```
Excluding rows with NA omits 167 rows.

# 1. How do the prediction variables interact to predict whether or not a client will sign on to a long-term deposit?

```{r}
sam = sample(3952,400)
train = bank_data[-sam,]
train = ovun.sample(y~., train,method="over", p=.5)$data
test = bank_data[sam,]
rf = randomForest(y ~ ., train)
pred = predict(rf, test)
pred = ifelse(pred == "yes", 1, 0)
test$y = ifelse(test$y=="yes", 1, 0)
mean(pred == test$y)
table(pred, test$y)
rf[["confusion"]]
```






